### 2023-08-15 state-resilience.tex
up

@@

We now consider the case $\Safe = \downarrow \Safe$ hence $\Bad = \uparrow \Bad$.
%
Unfortunately the {\sc State-resilience} problem is undecidable for WSTS in the case $\Safe = \downarrow \Safe$. As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in Appendix~\ref{HS-Minsk}.

>>>>

We are now interested in the 
%following
 undecidable \cite{araki1976PN} decision problem of {\em zero-reachability} in reset-VASS, which consists in, given a
reset-VASS $V=(Q,T)$, and $p(\textbf{u}) \in Q \times \N^d$,
deciding whether $\exists q \in Q ~ p(\textbf{u}) \to^* q(\textbf{0})$.
%
We reduce the zero-reachability problem to the problem of deciding whether a control-transition is live,
which we then reduce to {\sc State-resilience}.
A control-transition $t$ of a reset-VASS is {\em live} in a configuration $r(\textbf{w})$ if for each $q(\textbf{v}) \in \post^*(r(\textbf{w}))$ there exists a 
 $p(\textbf{u})$ with $q(\textbf{v}) \to^* p(\textbf{u})$ such that $t$ is enabled in $p(\textbf{u})$. We say the whole reset-VASS is live if all its control-transitions are
live. This leads to the following problem.

\problemx{$t$-liveness}
{reset-VASS $V=(Q,T)$, $t \in T$, initial configuration $s_0$}
{Is $t$ live in $s_0$ ? \\}

To better discuss $t$-liveness let us define more formally the set of configurations such that $t$ is enabled
$pre(t)=\{ p(\textbf{u}) \in S \mid \exists q(\textbf{v})$ such that $ p(\textbf{u}) \xrightarrow{t} q(\textbf{v}) \}$.

\begin{proposition}\label{reductions}
{\sc $t$-liveness} is reducible to {\sc State-resilience} in reset-VASS.
\end{proposition}

\begin{proof}
We reformulate $t$-liveness 
in a 
% WSTS $\mathscr{S}=(S,\rightarrow, \leq)$ with an initial state $s_0$ 
reset-VASS $(Q,T)$ 
with initial configuration $s_0$
as the following formula
\[ ~ \forall p(\textbf{u}) \in Q \times \N^d, 
~ s_0 \rightarrow^* p(\textbf{u}) \implies \exists q(\textbf{v}) \in U_t, ~ p(\textbf{u}) \rightarrow^{*} q(\textbf{v})\]
where
$U_t=\uparrow pre(t)$ 
is the upward closure of the preconditions to fire transition $t$.  
The problem reduces itself to {\sc State-resilience}
where $\Safe = U_t$.
\end{proof}

The following theorem is a transposition of Theorem~5.5 from \cite{peterson1981petri} from Petri nets to reset-VASS%, whose proof can be seen in Appendix~\ref{appendix}%
.
\begin{proposition}\label{liveness reset}
The  zero reachability problem can be reduced to the $t$-liveness problem in reset-VASS.
\end{proposition}

Since the zero-reachability problem for reset-VASS is undecidable, the reduction implies 
% the following:
%
%
{\sc %$ reset-VASS 
$t$-liveness} is undecidable.

Since {\sc $t$-liveness} is undecidable, from Proposition~\ref{reductions},  we deduce that {\sc State-resilience} is undecidable  for reset-VASS, which are WSTS with strong compatibility. Hence {\sc State-resilience} is undecidable  for WSTS with strong upward-compatibility. %This undecidability result furthermore implies the undecidability of the other two state resilience problems by straightforward reductions.

@@

We now consider the case $\Safe = \downarrow \Safe$ hence $\Bad = \uparrow \Bad$.
%
Unfortunately the {\sc State-resilience} problem is undecidable for WSTS in the case $\Safe = \downarrow \Safe$. As for resilience in WSTS when $\Safe = \downarrow \Safe$, it stems from undeciability of the corresponding problem in Minsky machines, the executions of which can be simulated by reset-VASS, as seen in % Appendix~\ref{HS-Minsk}.
